---
title: "BIOL606_hw17_Shah.rmd"
author: "Priyanshi Shah"
date: "2023-04-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vegan)
```


#CLASS
```{r}
veg <- read.table("bryceveg.txt", header = TRUE, row.names = 1)

sites <- read.table("brycesite.txt", header = TRUE, row.names= 1)

env = with(sites, cbind(elev, av, slope))

env.scale = as.data.frame(scale(env, center = TRUE, scale = TRUE))


```

#Switching from scaling 2 to 1 is giving us a diff Triplot depending upon the correlation between different predictor variables

#If the data is smaller, it would be nice to add the site names as well
#If the data is gigantic, then not adding the site names would be a good idea. 

#we can add a categorical predictor variable (as a factor)

#Predictor variables are the constraints on the ordination 

#weight and loading/ longer vector has more effect ------ PCA WORLD

#Final project 
STatistical analyses methods not how the data was collected 
File upload -> Midterm 3 

Same figures with the R code, no tweaking with adobe or picsart 

WHich package to cite?If the package is not doing anythig to recreate and manipulation 

db-rda, ypu're gonna use vegan, so cite that
if any bioinfo package 

ggplot2, ggfortify do not cite 

A lot, a lot of commenting 

Break your codes into chunks 

e.g file loading, scaling, plotting, analysis

No lines of code in the Word Document

```{r SNAILS DATA}
snails = read.table('SmallSnailData.txt', header = TRUE, sep = ' ')

shelldata = snails[,2:4]

rownames(shelldata) = as.character(snails$Site)

shell.center = as.data.frame(scale(shelldata))

#Subsetting site data

sitedata = snails[,5:8]

rownames(sitedata) = as.character(snails$Site)

site.center = as.data.frame(scale(sitedata))

```


```{r}




```


