---
title: "BIOL606_hw_Lecture06b_Shah.rmd"
output: html_document
date: "2023-02-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(ggfortify)
```
#Lecture
```{r}
cdata <- read.csv("lecture5_cucumberdamage.csv")
cdata
str(cdata) #gives the structure of the dataset
glimpse(cdata) #gives a more thorough structure

#Hypothesis: The total number of fruits produced is dependent on the herbivore damage to the leaves (i.e., "leafdamage") and hand pollination treatments

lm_model1 <- lm (totalfruits ~ leafdamage * pollination, data = cdata) #interactive model 
lm_model1

autoplot(lm_model1)

poisson_model <- glm(totalfruits ~ leafdamage * pollination, data = cdata, family = poisson) #look at the residual deviance, which is why we would use quasipoisson plots
poisson_model

anova(poisson_model)
anova(poisson_model, test = "Chi")

summary(poisson_model)

q_model <- glm(totalfruits ~ leafdamage * pollination, data = cdata, family = quasipoisson)
q_model

anova(q_model, test = "F")

q_model2 <- glm(totalfruits ~ pollination * leafdamage, data = cdata, family = quasipoisson) #changing the order of the variables. The order in which you enter the variables really matter. 
q_model2

anova(q_model2, test = "F")


#balanced design is when each category has same amounts, 
#unbalanced is when you don't have a lot of data. This is very common. The way you had barely any data for reptiles, amphibians!

```
#HOMEWORK

1. Import lecture5_NEWcucumberdamage.csv – NOTE THIS IS A NEW, UPDATED FILE.

```{r}
last_cucu <- read.csv("lecture5_NEWcucumberdamage.csv") #imported cucumber data FOR THE LAST TIME
glimpse(last_cucu) #looking at the data
```

2. If you're going to analyze totalfruits again, is this dataset balanced with regards to pollination & herbivory? What about pollination & leaf damage? 
3. Pick one of these pairs of IVs (pollination & herbivory, or pollination & leaf damage). Determine if these variables affect totalfruits, including an interaction, and going through the same steps you did last time to fit the model.

```{r}
#if the dataset was balanced, changing the order of variables won't change the final results, so let's see. 
#thank you Devon for this idea
checking1 <- glm(totalfruits ~ herbivory * pollination, data = last_cucu, family = quasipoisson)
checking1

checking2 <- glm(totalfruits ~ pollination * herbivory, data = last_cucu, family = quasipoisson)
checking2

anova(checking1) #type 1 (lowercase a) 
anova(checking2)

#the results of both these anova tests vary, which is an indication of an unbalanced dataset with regards to pollination and herbivory

checking3 <- glm(totalfruits ~ leafdamage * pollination, data = last_cucu, family = quasipoisson)
checking3

checking4 <- glm(totalfruits ~ pollination * leafdamage, data = last_cucu, family = quasipoisson)
checking4

anova(checking3) #type 1 (lowercase a) 
anova(checking4)

#the results of both these anova tests vary, which is an indication of an unbalanced dataset with regards to pollination and leafdamage
```

4. Test the variables in your model using both type 1 and type 3 approaches. Do the results change? Do the conclusions change?

```{r}
#TYPE 1 APPROACH for herbivory and pollination

checking1 <- glm(totalfruits ~ herbivory * pollination, data = last_cucu, family = quasipoisson)
checking1

checking2 <- glm(totalfruits ~ pollination * herbivory, data = last_cucu, family = quasipoisson)
checking2

anova(checking1, test = "Chisq")
anova(checking2, test = "Chisq")

#TYPE 1 APPROACH for leafdamage and pollination

checking3 <- glm(totalfruits ~ leafdamage * pollination, data = last_cucu, family = quasipoisson)
checking3

checking4 <- glm(totalfruits ~ pollination * leafdamage, data = last_cucu, family = quasipoisson)
checking4

anova(checking3, test = "Chisq") #type 1 (lowercase a) 
anova(checking4, test = "Chisq")


#TYPE 3 APPROACH for herbivory and pollination

checking1 <- glm(totalfruits ~ herbivory * pollination, data = last_cucu, family = quasipoisson)
checking1

checking2 <- glm(totalfruits ~ pollination * herbivory, data = last_cucu, family = quasipoisson)
checking2

library("car")
options(contrasts =c("contr.sum","contr.poly"))
checking1 <- glm(totalfruits ~ herbivory * pollination, data = last_cucu, family = quasipoisson)
checking2 <- glm(totalfruits ~ pollination * herbivory, data = last_cucu, family = quasipoisson)
Anova(checking1, type=3) #type 1 (lowercase a) 
Anova(checking2, type=3)

options(contrasts =c("contr.treatment","contr.poly"))


#TYPE 3 APPROACH for leafdamage and pollination

library("car")
options(contrasts =c("contr.sum","contr.poly"))
checking3 <- glm(totalfruits ~ herbivory * pollination, data = last_cucu, family = quasipoisson)
checking4 <- glm(totalfruits ~ pollination * herbivory, data = last_cucu, family = quasipoisson)
Anova(checking3, type=3) 
Anova(checking4, type=3)

options(contrasts =c("contr.treatment","contr.poly"))

#In comparison with the type I results, the significance values have changed. Using Type-1, the significance of deviation for herbivory was 2.274e-08, whereas using Type-3 it was different.  

```

